### 2016-05-30 capability-machine.tex
Changed V relation for execute cap to require to be able to execute from any address in range
added FTLR
Added environment/register relation

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{thmtools}
\usepackage{stmaryrd}
\usepackage{natbib}
\usepackage{url}

\usepackage{tikz}

@@

@@

%Assembly labels
\newcommand{\codelabel}[1]{\mathit{#1}}
\newcommand{\init}{\codelabel{init}}
\newcommand{\malloc}{\codelabel{malloc}}

@@

%LR
\newcommand{\intr}[2]{\mathcal{#1}\sem{#2}}
\newcommand{\valueintr}[1]{\intr{V}{#1}}
\newcommand{\exprintr}[1]{\intr{E}{#1}}
\newcommand{\contintr}[1]{\intr{K}{#1}}
\newcommand{\stdvr}{\valueintr{\asmType}}
\newcommand{\stder}{\exprintr{\asmType}}
\newcommand{\stdkr}{\contintr{\asmType}}
\newcommand{\observations}{\mathcal{O}}
\newcommand{\npair}[2][n]{\left(#1,#2 \right)}

@@

@@

@@

    \init: & \move{\refreg{r_1}}{\size} \\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{3} \\
           & \jmp{\malloc} \\

@@

           & \store{\refheap{r_1}}{(\encode(i_1))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \store{\refheap{r_1}}{(\encode(i_2))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \vdots \\
           & \store{\refheap{r_1}}{(\encode(i_7))} \\
           & \lea{\refreg{r_1}}{1} \\

@@

           & \move{\refreg{r_0}}{\refreg{r_1}} \\
           & \lea{\refreg{r_0}}{1} \\
           & \store{\refheap{r_1}}{\refreg{r_0}} \\
           & \lea{\refreg{r_1}}{1} \\

@@

           & \move{\refreg{r_3}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_3}}{17} \\
           & \store{\refheap{r_3}}{r_0} \\
           & \move{\refreg{r_3}}{0} \\

@@

           & \store{\refheap{r_1}}{0} \\
% Sent pointer to start of ticket dispenser code
           & \lea{\refreg{r_1}}{-8} \\

@@

           & \store{\refheap{\rio}}{-1} \\

@@

           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{9} \\
           & \store{\refheap{r_0}}{\refreg{\rio}} \\

@@

           & \move{\refreg{\rio}}{0} \\

@@

           & \restrict{\refreg{r_1}}{\refreg{r_1}}{(\encodePerm{(\entry)})} \\

@@

           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{5} \\
           & \restrict{\refreg{r_0}}{\refreg{r_0}}{(\encodePerm{(\entry)})} \\

@@

           & \jmp{\codelabel{adv}} \\
           & 0 \quad (\text{this address will be overwritten})\\
           & 0 \quad (\text{this address will be overwritten})\\

@@

           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{r_0}{-2} \\
           & \load{\refreg{r_1}}{\refheap{r_0}} \\

@@

           & \lea{r_0}{1} \\
           & \load{\refreg{r_2}}{\refheap{r_0}} \\

@@

           & \load{\refreg{r_0}}{\refheap{r_2}} \\

@@

           & \store{\refheap{r_1}}{\refreg{r_0}} \\

@@

Here $\size$ is 9. $i_1 \dots i_7$ refers to the instructions in the following program:

@@

 Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\exec,\init,\init+56,\init)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.

@@

                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}

@@

                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\stdcap[\exec]), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}

@@

@@

@@

                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}

@@

Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
\begin{align*}
  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\%TODO: What should this be?
  \observations (W) \defeq & \{ (\reg,\heap) \mid (\reg,\heap) \diverge \}
\end{align*}
\begin{align*}
  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
  \stdkr(W) \defeq & \{ \npair{c} \mid \\
                   & \quad \forall v\ldotp \in \stdvr(W) \Rightarrow \\
                   & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
                     \forall \heapSat[\heap]{n'}{W'}, \npair{v_1 \dots v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad ([\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \in \observations(W) \}
\end{align*}

\begin{align*}
  \stder \quad : & \quad \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
  \stder(W) \defeq & \{ \npair{\reg} \mid \\
                   & \quad \forall \npair{c} \in \stdkr(W) \\
                   & \qquad  \forall \heapSat[\heap]{n}{W}, \heap_f \ldotp \\
                   & \qquad \quad (\reg\update{r_0}{c},\heap \uplus \heap_f) \in \observations(W) \}
\end{align*}

\begin{definition}[Heap satisfaction/erasure]
\begin{align*}
  \heapSat{n}{W} & & \text{iff} & &
                                      \arraycolsep=0pt
                                      \begin{array}{l}
                                      \exists r: \dom(W) \rightarrow \Heaps \ldotp \\
                                      \qquad\memheap = \biguplus_{\iota\in \dom(W)} r(\iota)\\
                                      \quad\text{and}\\
                                      \qquad\forall \iota \in \dom(W) \ldotp \npair{r(\iota)} \in W(\iota)(W)
                                      \end{array}\\
\end{align*}
\end{definition}
\begin{lemma}[FTLR]
  For some configuration 
\end{lemma}

>>>>

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{thmtools}
\usepackage{stmaryrd}
\usepackage{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}

\usepackage{tikz}
\usepackage{multirow}

@@

\newcommand{\forcenewline}{$\phantom{v}$\\}

@@

%Assembly labels
\newcommand{\codelabel}[1]{\mathit{#1}}
\newcommand{\init}{\codelabel{init}}
\newcommand{\malloc}{\codelabel{malloc}}
\newcommand{\counter}{\codelabel{counter}}
\newcommand{\iocap}{\codelabel{iocap}}

@@

%LR
\newcommand{\intr}[2]{\mathcal{#1}\sem{#2}}
\newcommand{\valueintr}[1]{\intr{V}{#1}}
\newcommand{\exprintr}[1]{\intr{E}{#1}}
\newcommand{\contintr}[1]{\intr{K}{#1}}
\newcommand{\envintr}[1]{\intr{G}{#1}}
\newcommand{\stdvr}{\valueintr{\asmType}}
\newcommand{\stder}{\exprintr{\asmType}}
\newcommand{\stdgr}{\envintr{\asmType}}
\newcommand{\stdkr}{\contintr{\asmType}}
\newcommand{\observations}{\mathcal{O}}
\newcommand{\npair}[2][n]{\left(#1,#2 \right)}

@@

\newcommand{\mrlen}{5cm}
\newcommand{\bracket}[1]{\multirow{#1}{*}{\ensuremath{
 \left . \vphantom{\begin{array}{l}
 \ifthenelse{\equal{#1}{1}}{3\\}{
    \ifthenelse{\equal{#1}{2}}{3\\3\\}{
    \ifthenelse{\equal{#1}{3}}{3\\3\\3\\}{
    \ifthenelse{\equal{#1}{4}}{3\\3\\3\\3\\}{
    \ifthenelse{\equal{#1}{5}}{3\\3\\3\\3\\3\\}{
    \ifthenelse{\equal{#1}{6}}{3\\3\\3\\3\\3\\3\\}{
      3\\3\\3\\3\\3\\3\\3\\ %7
  }}}}}}
  \end{array}} \right \}}}
}
\newcommand{\annotate}[2]{\multirow{#1}{\mrlen}{\scriptsize #2}}

@@

  \begin{array}{r l l p{\mrlen}}

@@

    \init: & \move{\refreg{r_1}}{\size} &  \bracket{4} & \annotate{4}{Allocate memory for the ticket dispenser program (including memory for counter and capability for counter)}\\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} & & \\
           & \lea{\refreg{r_0}}{3} & & \\
           & \jmp{\malloc} & & \\

@@

           & \store{\refheap{r_1}}{(\encode(i_1))} & \bracket{7} & \annotate{7}{Store the ticket dispenser program in the newly allocated memory} \\
           & \lea{\refreg{r_1}}{1} & & \\
           & \store{\refheap{r_1}}{(\encode(i_2))} & & \\
           & \lea{\refreg{r_1}}{1} & & \\
           & \vdots & & \\
           & \store{\refheap{r_1}}{(\encode(i_7))} & & \\
           & \lea{\refreg{r_1}}{1} & & \\

@@

           & \move{\refreg{r_0}}{\refreg{r_1}} & \bracket{4} & \annotate{4}{Store a capability for the counter address after the ticket dispenser program} \\
           & \lea{\refreg{r_0}}{1} & & \\
           & \store{\refheap{r_1}}{\refreg{r_0}} & & \\
           & \lea{\refreg{r_1}}{1} & & \\

@@

           & \move{\refreg{r_2}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Store a capability for the counter in this code} \\
           & \lea{\refreg{r_2}}{\counter} & & \\
           & \store{\refheap{r_2}}{r_0} & & \\
           & \move{\refreg{r_2}}{0} & \bracket{1} & \annotate{1}{Overwrite capability for counter} \\

@@

           & \store{\refheap{r_1}}{0} & \bracket{1} & \annotate{1}{Initialize counter to 0} \\
% Change pointer to point to start of ticket dispenser code
           & \lea{\refreg{r_1}}{-8} & \bracket{1} & \annotate{1}{Capability points to start of td code \footnote{This is also the argument for the adversary}} \\

@@

           & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\

@@

           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Store the io capability on the stack} \\
           & \lea{\refreg{r_0}}{\iocap} & & \\
           & \store{\refheap{r_0}}{\refreg{\rio}} & & \\

@@

           & \move{\refreg{\rio}}{0} & \bracket{1} & \annotate{1}{Overwrite io register} \\

@@

           & \restrict{\refreg{r_1}}{\refreg{r_1}}{(\encodePerm{(\entry)})} &\bracket{1} & \annotate{1}{Restrict cap. to td} \\

@@

           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Setup a return pointer for the adversary} \\
           & \lea{\refreg{r_0}}{5} & & \\
           & \restrict{\refreg{r_0}}{\refreg{r_0}}{(\encodePerm{(\entry)})} & & \\

@@

           & \jmp{\codelabel{adv}} & \bracket{1} & \annotate{1}{Jump to the adversary} \\
           & 0 & \bracket{1} & \annotate{1}{Address reserved for io cap.} \\
           & 0 & \bracket{1} & \annotate{1}{Address reserved for counter cap.} \\

@@

           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capabiltiy} \\
           & \lea{r_0}{-2} & & \\
           & \load{\refreg{\rio}}{\refheap{r_0}} & & \\

@@

           & \lea{r_0}{1} & \bracket{2} & \annotate{2}{Retrieve the counter capability} \\
           & \load{\refreg{r_1}}{\refheap{r_0}} & & \\

@@

           & \load{\refreg{r_0}}{\refheap{r_1}} & \bracket{2} & \annotate{2}{Read the counter and write it to io address} \\

@@

           & \store{\refheap{\rio}}{\refreg{r_0}} & & \\

@@

Here $\size$ is 9. The variables $\counter$ and $\iocap$ are resepctively the offsets to the addresses reserved for the counter and io capabilities. $i_1 \dots i_7$ refers to the instructions in the following program:

@@

 Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+56,\init)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.

@@

                   & \quad (\forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W')) \}

@@

                   & \qquad \quad \npair[n']{[\pcreg \mapsto \stdcap[\exec], r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}

@@

                   & \{ \npair{\stdcap[\rwx] = c} \mid \\

@@

                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \\

@@

                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}

@@

Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
\begin{align*}
  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\%TODO: What should this be?
  \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid (\reg,\heap) \diverge \}
\end{align*}
\begin{align*}
  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
  \stdkr(W) \defeq & \{ \npair{c} \mid \\
                   & \quad \forall v\ldotp \in \stdvr(W) \Rightarrow \\
                   & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
                     \forall \heapSat[\heap]{n'}{W'}, \npair{v_1 \dots v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad (n,[\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \in \observations(W) \}
\end{align*}

\begin{align*}
  \stder \quad : & \quad \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
  \stder(W) \defeq & \{ \npair{\reg} \mid \\
                   & \quad \forall \npair{c} \in \stdkr(W) \ldotp \\
                   & \qquad  \forall \heapSat[\heap]{n}{W}, \heap_f \ldotp \\
                   & \qquad \quad (n,\reg\update{r_0}{c},\heap \uplus \heap_f) \in \observations(W) \}
\end{align*}

\begin{align*}
  \stdgr \quad : & \quad \Worlds \monfun \UPred{\Regs} \\
  \stdgr(W) \defeq & \{ \npair{\reg} \mid \\
                    & \quad \forall r \in \RegName \ldotp \\
                    & \qquad  \npair{\reg(r)} \in \stdvr(W) \}
\end{align*}

\begin{definition}[Heap satisfaction/erasure]
\begin{align*}
  \heapSat{n}{W} & & \text{iff} & &
                                      \arraycolsep=0pt
                                      \begin{array}{l}
                                      \exists r: \dom(W) \rightarrow \Heaps \ldotp \\
                                      \qquad\memheap = \biguplus_{\iota\in \dom(W)} r(\iota)\\
                                      \quad\text{and}\\
                                      \qquad\forall \iota \in \dom(W) \ldotp \npair{r(\iota)} \in W(\iota)(W)
                                      \end{array}\\
\end{align*}

\end{definition}
\begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
  If for all $n \in \nats$, $W \in \Worlds$, and register files $\reg$, $\npair{\reg} \in \stdgr(W)$, then $(n,\reg) \in \stder(W)$
\end{lemma}
